---
title: "Project 1"
author: "Manu"
date: "25 8 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Load data and packages

As a first step, I load the data and the required packages.

```{r load}
library(gridExtra)
library(ggplot2)

if (!file.exists("activity.csv") )
    {
     url <- 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'  
     download.file(url,destfile='repdata%2Fdata%2Factivity.zip',mode='wb')  
     unzip('repdata%2Fdata%2Factivity.zip')
    }
df <- read.csv("activity.csv")  
```

Now the data is ready and each question from the assignment can be answered step by step.

## 2. Histogram of the total number of steps taken each day

```{r total}
total <- tapply(df$steps, df$date, sum, na.rm=T)
head(total)
hist(total, xlab = "steps per day", main = "Total steps per day")
```

## 3. Mean and median number of steps taken each day

```{r mm}
tmean <- mean(total)
tmedian <- median(total)
tmean
tmedian
```


## 4. Time series plot of the average number of steps taken

```{r ts}
tstotal <- tapply(df$steps, df$interval, mean, na.rm=T)
plot(tstotal ~ unique(df$interval), type="l", xlab = "5 min intervals")
```


## 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r max}
max5 <- tstotal[which.max(tstotal)]
max5
```

## 6. Code to describe and show a strategy for imputing missing data

```{r missing1}
summary(df)
table(is.na(df) == TRUE)
```
For 2304 observations the steps are missing.

```{r missing2}
dfcomplete <- df
for (i in 1:nrow(dfcomplete)){
    if(is.na(dfcomplete$steps[i])){
        dfcomplete$steps[i]<- tstotal[[as.character(dfcomplete[i, "interval"])]]
    }
}
```

Missing values are replaced with the corresponding mean value from the time series object. 

## 7. Histogram of the total number of steps taken each day after missing values are imputed

```{r total2}
total2 <- tapply(dfcomplete$steps, dfcomplete$date, sum, na.rm=T)
head(total2)
hist(total2, xlab = "steps per day", main = "Total steps per day complete")
```

## 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends


```{r functions}
is_weekday <- function(z) {
  if (inherits(z, "zoo")) 
    z <- time(z)
  z <- as.POSIXlt(z)$wday
  z > 0L & z < 6L
}


is_weekend <- function(z) {
  if (inherits(z, "zoo")) 
    z <- time(z)
  z <- as.POSIXlt(z)$wday
  z < 1L | z > 5L
}
dfcomplete$weekend <- factor(is_weekend(dfcomplete$date), levels = c(TRUE, FALSE), labels = c("yes", "no"))

weekend <- subset(dfcomplete, weekend == "yes")
weekday <- subset(dfcomplete, weekend == "no")

summary(weekend)
summary(weekday)
```

```{r plotweekend}
tsweekend <- tapply(weekend$steps, weekend$interval, mean, na.rm=T)
```

```{r plotweekday}
tsweekday <- tapply(weekday$steps, weekday$interval, mean, na.rm=T)
```

```{r plotboth}
par(mfrow=2:1)
plot(tsweekend ~ unique(weekend$interval), type="l", xlab = "5 min intervals")
plot(tsweekday ~ unique(weekday$interval), type="l", xlab = "5 min intervals")
```
